---
title: "Lab 8 Homework"
author: "Derrick Tran"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
``` 

## Install `here`
The package `here` is a nice option for keeping directories clear when loading files. I will demonstrate below and let you decide if it is something you want to use.  
```{r}
#install.packages("here")
```

## Data
For this homework, we will use a data set compiled by the Office of Environment and Heritage in New South Whales, Australia. It contains the enterococci counts in water samples obtained from Sydney beaches as part of the Beachwatch Water Quality Program. Enterococci are bacteria common in the intestines of mammals; they are rarely present in clean water. So, enterococci values are a measurement of pollution. `cfu` stands for colony forming units and measures the number of viable bacteria in a sample [cfu](https://en.wikipedia.org/wiki/Colony-forming_unit).   

This homework loosely follows the tutorial of [R Ladies Sydney](https://rladiessydney.org/). If you get stuck, check it out!  

1. Start by loading the data `sydneybeaches`. Do some exploratory analysis to get an idea of the data structure.
```{r}
sydneybeaches<-readr::read_csv("data/sydneybeaches.csv")%>%
  clean_names()

sydneybeaches
```
```{r}
#janitor::clean_names("data/sydneybeaches.csv")
```

If you want to try `here`, first notice the output when you load the `here` library. It gives you information on the current working directory. You can then use it to easily and intuitively load files.
```{r}
#library(here)
```

The quotes show the folder structure from the root directory.
```{r}
#sydneybeaches <-read_csv(here("lab8", "data", "sydneybeaches.csv")) %>% janitor::clean_names()
```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

2. Are these data "tidy" per the definitions of the tidyverse? How do you know? Are they in wide or long format?

```{r}
skimr::skim(sydneybeaches)
```
```{r}
summary(sydneybeaches)
```
#in general, after using the clean names function the data looks somewhat tiddy howeverit does look troublesome to extract data from this table. from the Skimr function it doesnt look like theres any weird values in here, usually they would stick out here so this data is tidy enough I think
</div>

3. We are only interested in the variables site, date, and enterococci_cfu_100ml. Make a new object focused on these variables only. Name the object `sydneybeaches_long`
```{r}
sydneybeaches_long<-sydneybeaches%>%
  select(site,date,enterococci_cfu_100ml)

sydneybeaches_long
```


4. Pivot the data such that the dates are column names and each beach only appears once. Name the object `sydneybeaches_wide`
```{r}
sydneybeaches_wide<-sydneybeaches_long%>%
  pivot_wider(names_from = "date",
              values_from = "enterococci_cfu_100ml")

sydneybeaches_wide
```


5. Pivot the data back so that the dates are data and not column names.
```{r}
sydneybeaches_wide%>%
  pivot_longer(-site,
               names_to = "date",
               values_to = "enterococci_cfu_100ml")

```


6. We haven't dealt much with dates yet, but separate the date into columns day, month, and year. Do this on the `sydneybeaches_long` data.
```{r}
sydneybeaches_dates<-sydneybeaches_long%>%
  separate(date, into=c("day","month","year"), sep = "/")

sydneybeaches_dates
```


7. What is the average `enterococci_cfu_100ml` by year for each beach. Think about which data you will use- long or wide.
#so im not sure yet which format to use first, but lets try using long format first, based on the in class lab it seems long format makes it easier for us to play with the data.

```{r}
sydneybeaches_dates%>%
  select(site,year,enterococci_cfu_100ml)%>%
  group_by(year,site)%>%
  summarise(entero_avg=mean(enterococci_cfu_100ml,na.rm=T))
```


8. Make the output from question 7 easier to read by pivoting it to wide format.

```{r}
sydneybeaches_dates%>%
  select(site,year,enterococci_cfu_100ml)%>%
  group_by(year,site)%>%
  summarise(entero_avg=mean(enterococci_cfu_100ml,na.rm=T))%>%
  pivot_wider(names_from = "site",
              values_from = "entero_avg")
```

9. What was the most polluted beach in 2018?
```{r}
sydneybeaches_dates%>%
  select(site,year,enterococci_cfu_100ml)%>%
  group_by(year,site)%>%
  summarise(entero_avg=mean(enterococci_cfu_100ml,na.rm=T))%>%
  pivot_wider(names_from = "site",
              values_from = "entero_avg")%>%
  filter(year=="2018")
```

```{r}
sydneybeaches_dates%>%
  select(site,year,enterococci_cfu_100ml)%>%
  group_by(year,site)%>%
  summarise(entero_avg=mean(enterococci_cfu_100ml,na.rm=T))%>%
  filter(year==2018)%>%
  arrange(desc(entero_avg))
```

10. Perform one analysis of your choice on the data.
what are the dirtiest beaches in that its less than 10mL in entero concentration for 2018. its not the most informative but you could technically call this an analysis of the dirtiest clean beaches in 2018

```{r}
sydneybeaches_dates%>%
  select(site,year,enterococci_cfu_100ml)%>%
  filter(enterococci_cfu_100ml<10 &enterococci_cfu_100ml>0,year==2018)%>%
  group_by(site)%>%
  summarise(avg_entero=mean(enterococci_cfu_100ml,na.rm=T))%>%
  arrange(desc(avg_entero))
  


  
```


## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   