---
title: "Lab 3 Homework"
author: "Please Add Your Name Here"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Mammals Sleep
1. For this assignment, we are going to use built-in data on mammal sleep patterns. From which publication are these data taken from? Since the data are built-in you can use the help function in R.
```{r}
mammal_sleep <- msleep
```

2. Store these data into a new data frame `sleep`.
```{r}
sleep <- data.frame(mammal_sleep)
```

3. What are the dimensions of this data frame (variables and observations)? How do you know? Please show the *code* that you used to determine this below.  
```{r}
glimpse(sleep)
```
```{r}
structure(sleep)
```
```{r}
dim(sleep)
```

I did all 3 so we can see the numerical dimension and how it looks as a data frame because I figured they can be informative


4. Are there any NAs in the data? How did you determine this? Please show your code.  
```{r}
any(is.na(sleep) == TRUE, arr.ind=TRUE)
# i found this code on stack overflow. it tells me there is NA data entries but not where exactly 
is.na.rm=T
```


5. Show a list of the column names is this data frame.
```{r}
names(sleep)
```

6. How many herbivores are represented in the data? 
There are 32 herbivores. it look like there were some hints given on discord line 66 is from a classmate. it seems more efficient but I included mine anyways
```{r}
herbivores <- subset(sleep, vore == "herbi" )
herbivores
```


```{r}
table(sleep$vore)
```

7. We are interested in two groups; small and large mammals. Let's define small as less than or equal to 1kg body weight and large as greater than or equal to 200kg body weight. Make two new dataframes (large and small) based on these parameters.
```{r}
#structure(sleep$bodywt)
small_animal <- subset(sleep, bodywt <= 1)
small_animal
```

```{r}
large_animal <- subset(sleep, bodywt >= 200)
large_animal
```


8. What is the mean weight for both the small and large mammals?
the mean body weight for small animals is 0.26 and 1747 for large animals
```{r}
mean(small_animal$bodywt)
```

```{r}
mean(large_animal$bodywt)
```

9. Using a similar approach as above, do large or small animals sleep longer on average?  
larger animals tend to sleep longer then smaller animals 
```{r}
mean(small_animal$sleep_total)
```
```{r}
mean(large_animal$sleep_total)
```
small animals sleep more on average then large animals

10. Which animal is the sleepiest among the entire dataframe?

I found this on stack overflow, i think this allows you to take a 'slice' of the row of the specified condition that you want, in this case i wanted to use the data frame sleep and to take a slice of the row with the minimum sleep total, also i wasnt sure if the least amount of sleep is the sleepiest or the most amount of sleep so i included both
```{r}
sleep %>% slice_min(sleep_total)
```
```{r}
sleep %>% slice_max(sleep_total)
```

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   